---
title: "Growth models"
author: "Kristen Tuosto"
date: "2023-04-06"
output: html_document
---

#Goal: Model growth using quadratic log-log method
```{r}

library(dplyr)
library(ggplot2)
library(tidyverse)


#Read data into R
df <- read.csv("~Fake_growth_data.csv")

```

#Create new data frame with just required information (i.e., ID, Sex or Gender, Age, body measurement attempting to model
```{r}

filter_sample1 <- function(df, x) {
 x <- filter(df, Bone.Code == x) # Optional if your data is in long format, otherwise it can be dropped; to fit another data set 'Bone.Code' to the column title to be filtered from.
 x[x==""] <- NA
 x <- x %>% drop_na(ID, Age, MXL) #Replace with column titles for the information you would like to drop NAs 
 x$Sex <- as.factor(x$Sex)
 x$MXL <- as.numeric(x$MXL)
 x$Age <- as.numeric(x$Age)
 x <- data.frame(x$ID, x$Sex, x$Age, x$MXL, x$Age_Class)
 x <- x %>% rename("ID" = "x.ID",
                   "sex" = "x.Sex",
                   "age" = "x.Age",
                   "age_class" = "x.Age_Class",
                   "MXL" = "x.MXL")
 return(data.frame(x))
}

df_new <- filter_sample1(df, 'Tibia')
```

# Means, min, and max function by age class and sex

```{r}

# Filter the data frame by sex and age groups

female_AD <- subset(df_new, sex == "F" & class == "AD")
female_SAD <- subset(df_new, sex == "F" & class == "S-AD")
female_JUV <- subset(df_new, sex == "F" & class == "JUV")
female_INF <- subset(df_new, sex == "F" & class == "INF")
male_AD <- subset(df_new, sex == "M" & class == "AD")
male_SAD <- subset(df_new, sex == "M" & class == "S-AD")
male_JUV <- subset(df_new, sex == "M" & class == "JUV")
male_INF <- subset(df_new, sex == "M" & class == "INF")

# Define the mean_max_min function
mean_max_min <- function(df) {
  mean_vals <- mean(df$MXL)
  max_vals <- max(df$MXL)
  min_vals <- min(df$MXL)
  df_summary <- data.frame(
    mean = mean_vals,
    max = max_vals,
    min = min_vals
  )
  return(df_summary)
}

# Create a list of data frames
data_frames <- list(female_AD, female_SAD, female_JUV, female_INF, male_AD, male_SAD, male_JUV, male_INF )

# Apply the mean_max_min function to each data frame in the list
result_list <- lapply(data_frames, mean_max_min)

# Combine the summary data frames into a single data frame
result_df <- bind_rows(result_list)

# Add new column & combine with date frame
label <- c("female_AD", "female_SAD", "female_JUV", "female_INF", "male_AD", "male_SAD", "male_JUV", "male_INF")

result_df_label <- cbind(result_df, label)

#Relocation label column
new_result_df <- result_df_label %>% relocate(label)


# Print the final result
print(new_result_df)

# Write an excel sheet
library(writexl)
write_xlsx(new_result_df,"C:\\Users\\ktuos\\Desktop\\tibia_summary_results.xlsx")
```

#Basic loess plots just to visualize trends before modeling growth
```{r}
#Loess plots with combined sex
loess_plots_sex <- function(df, bm){ # bm needs to be df$bm
  plot <- ggplot(data= df, aes(x = age, y = bm, color = sex)) + geom_smooth(method = 'loess', formula = 'y ~ x') + geom_point(aes(color=sex), size = 3) + theme_classic() + labs(x="Age (years)") + scale_colour_discrete(labels = c("Female", "Male")) + labs(color = "Sex") 
  return(plot)
} 

#Loess plots sex separated 
loess_plots <- function(df, bm){ # bm needs to be df$bm
  plot <- ggplot(data= df, aes(x = age, y = bm)) + geom_smooth(method = 'loess', formula = 'y ~ x') + geom_point(size = 3) + theme_classic() + labs(x="Age (years)")
  return(plot)
} 

#Return plots can be further modified with ggplot codes

```

#Quadratic Log-Log regression 
Quadratic log-log regression - sex considered together
```{r}

loglog_sex <- function(df, dm) {
  loglog <- lm(log(dm) ~ poly(log(age), 2, raw = TRUE) + sex, data = df) 
  print(summary(loglog))
  print(plot(loglog))
  intercept_loglog <- loglog$coefficients[1]
  logage_loglog <- loglog$coefficients[2]
  logage2_loglog <- loglog$coefficients[3]
  
  plotlog <- function(x) {
    exp(intercept_loglog + log(x)*logage_loglog + log(x)^2*logage2_loglog)} #Function from Levy et al., 2022, plots quadratic log-log line
  
  plot <- ggplot(data = df, aes(x=age, y=dm)) + 
    geom_point(aes(color=sex), size = 3) + 
    stat_function(fun = plotlog) + 
    labs(x='Age (years)', title='log(y) ~ log(x) + log(x)^2') + 
    theme_classic() +      
    scale_x_continuous(n.breaks=10) + 
    scale_colour_discrete(labels = c("Female", "Male")) + 
    labs(color = "Sex")
    return(plot)
}
 
#Return plots can be further modified with ggplot codes
```

#Quadratic log-log regression for samples with sex separated and not considered 
```{r}

loglog <- function(df, dm) {
  loglog <- lm(log(dm) ~ poly(log(age), 2, raw = TRUE), data = df)
  print(summary(loglog))
  print(plot(loglog))
  intercept <- loglog$coefficients[1]
  logage <- loglog$coefficients[2]
  logage2 <- loglog$coefficients[3]
  
  plotlog <- function(x) {
    exp(intercept + log(x)*logage + log(x)^2*logage2)} #Function from Levy et al., 2022, plots quadratic log-log line
  
  plot <- ggplot(data = df, aes(x=age, y=dm)) + 
    geom_point(size = 3) + 
    stat_function(fun = plotlog) + 
    labs(x='Age (years)', title='log(y) ~ log(x) + log(x)^2') + 
    theme_classic() +      
    scale_x_continuous(n.breaks=10) 
    return(plot)
}

#Return plots can be further modified with ggplot codes

```

#Code test for sex combined functions
```{r}
loess_plots_sex(df_new, df_new$MXL)

tibiaplot <- loglog_sex(df_new, df_new$MXL)

tibiaplot + labs(y="MXL")
```
#Code test
```{r}
loess_plots(female_df, female_df$MXL)

loglog(female_df, female_df$MXL)
```







